---
title: "Plotly diamonds data"
author: "Alexey Maranda"
date: '27 February 2017 ã '
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(plotly)
```

## Goals
Create a web page presentation using R Markdown that features a plot created with Plotly.


## Slide with data description
We'll use a dataset containing the prices and other attributes of almost 54,000 diamonds. 
A data frame with 53940 rows and 10 variables:

- price - price in US dollars (\$326–\$18,823)

- carat - weight of the diamond (0.2–5.01)

- cut - quality of the cut (Fair, Good, Very Good, Premium, Ideal)

- color - diamond colour, from J (worst) to D (best)

- clarity - a measurement of how clear the diamond is (I1 (worst), SI1, SI2, VS1, VS2, VVS1, VVS2, IF (best))

## Slide with data exploration

```{r, echo = TRUE}

data(diamonds)
head(diamonds)

```

## Slide with Plot

```{r echo=TRUE , message = FALSE , warning = FALSE}
plot_ly(diamonds, x = ~clarity, color = ~cut ) 

```

